
\begin{acknowledgements}

% stop indents
{\parindent0pt

Primary acknowledgement goes to the traditional ownership of the lands and seas that host the Australian state; these coasts have been known for a long time.
\newline{}
    
Big thanks to Gary Brassington for enabling and encouraging this whole effort.
And to Kevin Walsh for guiding the University engagement. 


Many colleagues at the Bureau of Meteorology have provided support and motivation. 
James, Bill and Mike from the tidal unit are due a special mention.    
Discussions on sea level with James Taylor, Ben Hague, Simon Louis, David Grant and many more over the years have helped keep the focus on the practical implications for forecasting.
Work with Mikhail Entel, Diana Greenslade and Stefan Zieger have been particularly influential.
\newline{}
This study has of course only been possible thanks to the foundations built by the wider Bluelink community - a uniquely successful and persistent cross-organisation collaboration.
Also beyond the Bureau, thanks go to Zarina at the Australian Hydrographic Service and to Ed Zaron and Mike Foreman for early productive discussions. 
\newline{}
Final thanks to the ever-fantastic Kulja and all those non-digital mudflats and mangroves.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newline{}

Observational data shared with the Bureau from external agencies was important for this study:
\begin{itemize}
    \item Manly Hydraulics Laboratory %\url{http://mhl.nsw.gov.au/}
    \item Queensland Government Department of Environment, Land and Water
    \item Ports Australia
    %\url{https://www.qld.gov.au/environment/coasts-waterways/beach/storm-sites]}
\end{itemize}


% template from git bloke	
Credit for the LaTeX template to John Papandriopoulos.  %\url{http://jpap.org/projects.html}

}   % end no indent
\end{acknowledgements}
