%Section ----------------------------------------------------------------------------------------------------------------------------------------------------
\newpage
\section{Research plan}  
\label{S:work_plan} 


%--------------------------------------------------
\subsection{Overview}
%--------------------------------------------------

The body of this thesis will be approached as a number of discrete research problems; connected along a conceptual path branching from the preceding Literature Review [Section \ref{S:REVIEW}].\\
Several of the investigations will developed to form viable independent journal publications.  Progress towards a coherent thesis will be guided by the goal of successively creating a series of peer-reviewed publications.\\




An overview of the plan is provided by the following sequence of questions.   
% \begin{mdframed}[hidealllines=true,backgroundcolor=yellow!30]
\BoxBegin
\begin{enumerate}
\item Does coastal sea level from a 0.1 degree B-grid \OGCM{} have physical meaning, and with what limitations? 
	\par[pg \pageref{S:plan_whichcell}]
\item How is the surface height signature of coastally trapped waves represented in \BL{} and why? 
	\par[pg \pageref{S:plan_CTW}]
\item Is there any tidal component in the sea level forecast by this nominally nontidal system and why?
	\par[pg \pageref{S:plan_nontidal_tidal}]
\item How can standard harmonic tidal analysis be approached when the end-use is aggregation with dynamic models?
	\par[pg \pageref{S:plan_insitu_analysis}]
\item Can explicit resolution of tidal physics within and \OGCM{} realistically provide any benefit for coastal sea level?
	\par[pg \pageref{S:plan_OFAMR}]
\item Can a regional tidal model be optimised for use in connection with a tide resolving regional \OGCM{}?
	\par[pg \pageref{S:plan_OTIS}]
\item Is a tide resolving \OGCM{} sensitive to the details of tidal body forcing and can the formulation be improved?
	\par[pg \pageref{S:plan_bodyforcing}]
\item And finally, can the proceeding work inform an initial discussion of the implications of explicit tidal resolution for data assimilation?
	\par[pg \pageref{S:plan_DA}]
\end{enumerate}
%\end{mdframed}
\BoxEnd


%----------------------------------------------------------------------------------------------------------------------------------------------------
%----------------------------------------------------------------------------------------------------------------------------------------------------

\input{ataylor_phd_3.1_whichcell.tex}
\input{ataylor_phd_3.2_CTW.tex}
\input{ataylor_phd_3.3_tidal_nontidal.tex}
\input{ataylor_phd_3.4_harmonic_analysis.tex}
\input{ataylor_phd_3.5_explicit_tides.tex}
\input{ataylor_phd_3.6_tidal_BC.tex}
\input{ataylor_phd_3.7_body_force.tex}
\input{ataylor_phd_3.8_DA.tex}


















