%--------------
\chapter{Research questions}

The maturation of mesoscale ocean prediction systems including \BL{} has bought qualitatively new types of sea level prognosis into the same context in which traditional tide tables have long been produced.
Whilst the canonical numerical approach to coastal sea level involves the pursuit of simulations with ever-higher spatial resolutions, this existing setting motivates an investigation into signal representation and the value of larger and longer scales of predictability.

\BoxBegin
\em{What is at the intersection of conventional tidal analysis and mesoscale ocean prediction, and what are the ramifications for operational sea level forecasts?}
\BoxEnd

This desktop study addresses the question using data and forecast systems of the Australian Bureau of Meteorology. New methods for combining, evaluating and delivering nominally “tidal” sea level information are proposed and analysed; with a focus on representation issues and system compatibility.


Despite the relatively low spatial resolution of the coast can \BL{} skilfully forecast coastal sea level and with what limitations? 

How is the surface height signature of coastally trapped waves represented in operational forecast models? 

Can nominally non-tidal sea level forecast systems simulate tidal patterns and why?


How can harmonic tidal prediction be approached when one end-use is aggregation with dynamic models?


%--------------
\chapter{Introduction}

%--------------
% review
\include{chapters/intro-context}
\include{chapters/intro-operational}
\include{chapters/intro-tides}
\include{chapters/intro-overlap}






